---
title: "mon8c9a_RefPoints_10B0"
author: "Paz Sampedro"
date: "2025-02-28"
output:
  html_document:
  theme: flatly
toc: true
toc_float:
  collapsed: false
pdf_document:
  toc: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

# The ICES approach to setting Reference Points

This document outlines the steps involved in estimating Precautionary Approach and MSY reference points for white anglerfish in area 8c and 9a as part of the WKBSS3 benchmark 2025.

It follows the current technical guidelines "ICES fisheries management reference points for category 1 and 2 stocks" (ICES, 2022) and the results of WKNEWREF2 (ICES, 2024).

```{r}
library(r4ss)
library(ss3om)      
library(FLCore)
library(msy)        
library(icesAdvice)
library(ggplot2)
library(dplyr)
library(FLRef)
```

### Step 0. Prepare the data in FLR

```{r}
# Directory
dir <- "C:/Users/psampedro/Desktop/WKSS3_WORK/BenchmarkModelsv2/FINALMODEL_MON8C9A/Model3.Kest_Msex_PTNETSLOG"
# SS3 output
replist <- SSgetoutput(dirvec = dir, getcovar = F, verbose = FALSE)[[1]]

# FLR objects
mon.stk <- readFLSss3(dir)  # FLR
stk = mon.stk

mon.sr  <- readFLSRss3(dir) # FLSR
dnms <- dimnames(mon.stk)

# Characteristics
# Ages    
ages  <- as.numeric(dnms$age)
# Years   
yrs   <- as.numeric(dnms$year)
# Units   
units <- dnms$unit
# Seasons
ssons <- as.numeric(dnms$season)
# Fbar
mon.stk@range['minfbar'] <- 2
mon.stk@range['maxfbar'] <- 15

# Set some other slots so that SSB can be calculated (M is approximate as the weight are not estimated exactly at the spawning time)
m.spwn(mon.stk)[] <- 0
harvest.spwn(mon.stk)[] <- 0

m.spwn(stk)[] <- 0
harvest.spwn(stk)[] <- 0

# Collapse to one season and unit
stk <- nounit(mon.stk) # combine sexes
stk <- noseason(stk)   # combine seasons

# Recruitment correction (the second part of the object is necessary only if you have multiple settling periods which is not in this)
raf <-  stock.n(mon.stk)[1,,"F",1,] 
ram <-  stock.n(mon.stk)[1,,"M",1,] 
stock.n(stk)[1,] <- raf + ram

#raf <-  stock.n(mon.stk)[1,,"F1",1,] + stock.n(mon.stk)[1,,"F3",3,]
#ram <-  stock.n(mon.stk)[1,,"M1",1,] + stock.n(mon.stk)[1,,"M3",3,] 
#stock.n(stk)[1,] <- raf + ram

### Natural mortality correction
mf <- replist$M_at_age %>% filter(Sex == 1 & Yr %in% yrs) %>% select(Yr, '0') %>% 
  tidyr::pivot_wider(names_from = Yr, values_from = '0') %>% unlist()
mm <- replist$M_at_age %>% filter(Sex == 2 & Yr %in% yrs) %>% select(Yr, '0') %>% 
  tidyr::pivot_wider(names_from = Yr, values_from = '0') %>% unlist()
m(stk)[1,] <- (raf * mf + ram * mm)/(raf+ram) 

mat(stk)[1,] <- 0

# Correct harvest
harvest(stk) <-  harvest(stock.n(stk), catch=catch.n(stk), m=m(stk))
```

Comparison of SSB and recruitment values: FLR and SS3:
  
  SSB Plot

```{r}
SSssb <- subset(replist$derived_quants,substring(Label,1,6)%in%c('SSB_19','SSB_20')) %>% 
  mutate(Yr=as.numeric(substring(Label,5,9)))
plot(SSssb$Yr,SSssb$Value,xlab="Year",ylab="SSB", ylim=c(0,max(c(ssb(stk),SSssb$Value,unitSums(ssb(mon.stk[,,"F"]))))*1.05))
lines(stk@range[[4]]:stk@range[[5]],ssb(stk))
lines(mon.stk@range[[4]]:mon.stk@range[[5]],as.vector(unitSums(ssb(mon.stk[,,"F",1,1]))),col='red')
legend('topright',c('SS SSB','FLR both Sexes','FLR Female SSB'),pch=c(1,NA,NA),lty=c(NA,1,1),col=c(1,1,2))
```

Recruitment Plot

```{r}
SSrec <- subset( replist$derived_quants, substring(Label,1,7) %in% c('Recr_19','Recr_20')) %>%
  mutate(Yr=as.numeric(substring(Label,6,10)))

plot(SSrec$Yr,SSrec$Value, ylim = c(min(c(rec(stk),SSrec$Value))*0.95,max(c(rec(stk),SSrec$Value))*1.05),
     xlab="Year",ylab="Rec")
lines(stk@range[[4]]:stk@range[[5]],rec(stk))
legend('topleft',c('SS rec','FLR rec'),pch=c(1,NA),lty=c(NA,1),col=1)
```

```{r}
alpha <- c(SSssb$Value[1:(length(yrs))]/ssb(stk)[,,])
mat(stk) <- sweep(mat(stk), 2, alpha, "*")
```

## Step 1. Identifying appropriate data

### Plot Time series of Recruitment and SSB

```{r}
# Plot time series
plot(c(ssb(stk)),c(rec(stk)),cex=0,xlab='ssb',ylab='rec')
text(c(ssb(stk)),c(rec(stk)),substring(stk@range['minyear']:stk@range['maxyear'],3,4),cex=0.7)
```

Stock development over the full time series:
  
```{r}
plot(stk) + facet_wrap(~qname, scales="free") + theme_minimal()
```


The SR estimated by the Stock Synthesis model is the option to be used, to be in line with the model estimates (Fixed Beverton Holt).

```{r}
# SR para from SS
r0 <- replist$parameters$Value[replist$parameters$Label=="SR_LN(R0)"]
R0 <- exp(r0)
h <- replist$parameters$Value[replist$parameters$Label=="SR_BH_steep"]
B0 <- c(replist$derived_quants$Value[replist$derived_quants$Label=="SSB_Virgin"])
# other SR para are not used here
sigmaR <- replist$parameters$Value[replist$parameters$Label=="SR_sigmaR"]


FixedBevholt <- function (ab,ssb) log(4*h*R0*ssb/(B0*(1-h)+ssb*(5*h-1)))
```

Checking how SR looks like, also compare with BH fit to the data (without fixing h). FixedBevHolt is the assessment output.

```{r}
set.seed(1)
nsamp <- 1000 #increase number of samples for final run (e.g 1000)
fit_bh_temp <- eqsr_fit(stk, nsamp = nsamp, models = c("FixedBevholt","Bevholt")) 
fit_bh <- eqsr_fit(stk, nsamp = nsamp, models = c("FixedBevholt")) 
eqsr_plot(fit_bh_temp, ggPlot=T)
fit_bh$stk <- stk #Want to have data up to 2023 available for selectivity
```

## Step 2. Stock type

The stock is Type 5: Stocks showing no evidence of impaired recruitment or with no clear relation between stock and recruitment (no apparent SR signal).

The WKNEWREF 2024 (next ICES guidelines) recommends used the **Empirical Blim defined as the the average of the lowest three SSB-s that resulted in above median recruitments**.

## Step 3. Estimating biomass limit reference points

It was decided to define Blim as 10% of Virgin Biomass. It is supposed that is well calculated by the model.

#### 10% B0

```{r}
B0_10 <- 0.10*B0
print(B0_10)
Blim <- B0_10
```

```{r}
# Plot for comparison Blim options
eqsr_plot(fit_bh)
abline(v=0.10*B0,lty=3, col="darkgreen")
text(0.10*B0-300,3500,'10% B0',pos=3,cex=1,col='blue',srt=90)

```

#### Blim

Blim is 10% of B0.

## Step 4. Other PA reference points from Blim

### Bpa is Blim plus assessment error:

sigmaSSB

```{r}
sigmaSSB <- round(with(subset(replist$derived_quants,Label=='SSB_2021'),StdDev/Value),3)
print(sigmaSSB)
```

Bpa:
  
```{r}
# This is higher than 0.2, then we use it. sigmaSSB <- 0.2 # default
sigmaSSB <- max(sigmaSSB, 0.2)
Bpa <- round(Blim * exp(1.645 * sigmaSSB))
print(Bpa)
```

### Flim

It is no longer used as a reference point in the ICES framework. It is included here for completeness. The preferred method is simulating a stock with a segmented regression SR relationship, with the point of inflection at Blim, thus determining the F = Flim which, at equilibrium, yields a 50% probability of SSB \> Blim. Note that this simulation should be conducted based on a fixed F (i.e. without inclusion of a Btrigger) and without inclusion of assessment/advice errors. This means Btrigger, Fcv, and Fphi should all be set to zero

```{r}
sigmaR <- replist$sigma_R_info[1,'SD_of_devs']
recruit = replist$recruit
recyrs = recruit$Yr[recruit$era =="Main"]
maindevs = recruit[recruit$Yr%in%recyrs,'dev']
rho = acf(maindevs)$acf[2]
fit.segreg <- eqsr_fit(stk, nsamp = nsamp, models = c("Segreg"))
sim_segregBlim <- eqsim_run(fit.segreg,
                            Fcv=0, Fphi=0, SSBcv=0,
                            rhologRec=rho,
                            Btrigger = 0, Blim=Blim,Bpa=Bpa,
                            Nrun=200, Fscan = seq(0,2,len=100),verbose=F)

Flim <- round(sim_segregBlim$Refs2['catF','F50'],3)
print(Flim)
```

Flim is estimated at `r Flim`.

## Step 5. Fmsy and Btrigger

$F_{msy}$ should initially be calculated based on an evaluation with the inclusion of stochasticity in a population (i.e. recruitment, M, maturity, growth) and fishery (e.g. selectivity) as well as assessment/advice error. This is a constant F, which should provide maximum yield without biomass constraints (without MSY Btrigger). Error is included as this is the condition analogous to management strategy evaluations (MSEs) that will prevail in practice. Note that in order to ensure consistency between the precautionary and the MSY frameworks, $F_{msy}$ is not allowed to be above Fpa; therefore, if the $F_{msy}$ value calculated initially is above Fpa, $F_{msy}$ is reduced to Fpa.

```{r}

## Forecast error (see Guidance document for details on calculation of these values)

# F
cvF  <- 0.212                                 # Default = 0.212
phiF <-	0.423                                 # Default = 0.423

# SSB
cvSSB <- 0                                    # Default = 0
phiSSB <- 0                                   # Default = 0
fit_bh$sr.sto$cv <- sigmaR
sim_noTrig <- eqsim_run(fit_bh,
                        Fcv=cvF, Fphi=phiF, SSBcv=cvSSB,
                        Btrigger = 0, Blim=Blim,Bpa=Bpa,
                        Nrun=500, Fscan = seq(0,2,len=100),verbose=F)

Fmsy_tmp <- round(sim_noTrig$Refs2["lanF","medianMSY"],3)
eqsim_plot_range(sim_noTrig, type="median")
```

Fmsy is initially estimated as `r Fmsy_tmp`.

## Step 6. Bmsy5pc

MSY Btrigger should be selected to safeguard against an undesirable or unexpected low SSB when fishing at FMSY.

In this case the stock has been fished below Fmsy for \> 10 years so MSY Btrigger is set equal to Bmsy5pc, that reflects the 5th percentile definition of MSY Btrigger.

```{r}
data.05 <- sim_segregBlim$rbp
x.05 <- data.05[data.05$variable == "Spawning stock biomass", ]$Ftarget
b.05 <- data.05[data.05$variable == "Spawning stock biomass", ]$p05
i <- which(x.05 < (Flim+0.2))
b.lm <- loess(b.05[i] ~ x.05[i])
Bmsy5pc <- round(predict(b.lm,Fmsy_tmp))

plot(b.05~x.05, ylab="SSB", xlab="F")
abline(v=Fmsy_tmp) 
lines(x.05[i],c(predict(b.lm)),type='l')
abline(h=Bmsy5pc)
abline(h=Bpa,lty=3)
text(0,Bmsy5pc+400,'Bmsy5pc',pos=4)
text(0,Bpa+400,'Bpa',pos=4)
print(Bmsy5pc)
Btrigger <- max(Bpa,Bmsy5pc)
print(Btrigger)
```

We should use Btrigger = Bpa =`r formatC(Bpa,format='d',big.mark=",")` because Bpa is higher.

## Step 7. FMSY + MSY Btrigger combination

ICES MSY AR should be evaluated to check that the FMSY and MSY Btrigger combination fulfills the precautionary criterion of having less than 5% annual probability of SSB \< Blim in the long term. The evaluation must include realistic assessment/advice error and stochasticity in population biology and fishery selectivity.

```{r}
sim_Trig <- eqsim_run(fit_bh,
                      Fcv=cvF, Fphi=phiF, SSBcv=cvSSB,
                      Btrigger = Btrigger, Blim=Blim,Bpa=Bpa,
                      Nrun=500, Fscan = seq(0,2,len=101),verbose=F)
eqsim_plot(sim_Trig,catch=F)
eqsim_plot_range(sim_Trig, type="median")
Fp05 <- round(sim_Trig$Refs2["catF","F05"],3)
print(Fp05)
```

If the precautionary criterion evaluated in point 3 is not met, then FMSY should be reduced from the value calculated above until the precautionary criterion is met (i.e. reduce FMSY to FMSY = Fpa). Fpa is Fp05 and is estimated at `r Fp05`.

```{r}
Fmsy <- round(min(sim_noTrig$Refs2["lanF","medianMSY"],Fp05),3)
Fupper <- round(min(sim_noTrig$Refs2["lanF","Medupper"],Fp05),3)
Flower <- round(min(sim_noTrig$Refs2["lanF","Medlower"],Fp05),3)
```

### Stock Synthesis Reference Points

```{r}
BMSY <- replist$derived_quants$Value[replist$derived_quants$Label=="SSB_MSY"]
FMSY <- replist$derived_quants$Value[replist$derived_quants$Label=="annF_MSY"]
BMSY
FMSY
```

BMSY =`r formatC(BMSY)`
FMSY =`r formatC(FMSY)`

## Reference points - Table

The estimated reference points are shown below.

Reference Point|Value                              |Rationale
---------------|-----------------------------------|-----------------------------------------
Blim           |`r formatC(Blim,format='d',big.mark=",")`     |10% B0
Bpa            |`r formatC(Bpa,format='d',big.mark=",")`      |Blim with assessment error
MSY Btrigger   |`r formatC(Btrigger,format='d',big.mark=",")` |Bpa
Fpa            |`r Fp05`                                      |F with 95% probability of SSB ≥ Blim (BH with Btrigger)
Fmsy           |`r Fmsy`                                      |Stochastic simulations
FmsyLower      |`r Flower`                                    |Stochastic simulations
FmsyUpper      |`r Fupper`                                    |Fpa
Bmsy5pc        |`r formatC(Bmsy5pc,format='d',big.mark=",")`  |5% probability of SSB < Blim

Current stock status: F is below Fmsy and SSB is above Blim and Btrigger.

```{r}
plot(window(stk,start=1980) ) + facet_wrap(~qname,scales='free_y') + geom_hline(aes(yintercept=Fmsy),data=data.frame(qname='F'),lty=3, col="red") + geom_hline(aes(yintercept=Blim),data=data.frame(qname='SSB'),lty=3, col="red") + geom_hline(aes(yintercept=Btrigger),data=data.frame(qname='SSB'),lty=2, col="orange") +theme_light() 
```
