---
title: "Pollack 6,7 Eqsim"
author: "Hans Gerritsen, modified by Tim Earl and Max Cardinale"
date: "19/02/2025"
output: pdf_document
---

## The ICES approach to setting Reference Points

This Markdown document outlines the steps involved in estimating PA and MSY reference points for pollack in divisions 6 and 7 as part of WKBSS3 benchmark 2025. It follows the current technical guidelines: <https://www.ices.dk/sites/pub/Publication%20Reports/Guidelines%20and%20Policies/16.04.03.01_Reference_points_for_category_1_and_2.pdf>.

**Some plot legends refer to landings or catch - These should all be read at "Total removals" - Commercial landings, recreational landings + the proportion of recreational releases assumed to die following release**


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,message=FALSE, warning=FALSE)
options(scipen=999)
```

## Get the data in FLR
```{r}
library(r4ss)
library(ss3om) #for readLFSss3
library(FLCore)
library(msy)
library(icesAdvice)
library(ggplot2)
library(dplyr)

dir <- getwd()
ass.wd=file.path(dir, "../",sep="")

replist <- SS_output(paste0(ass.wd),verbose=F,printstats=F)
stock0 <- readFLSss3(paste0(ass.wd))

stock0@harvest@units <- 'f'

# check fbar range used in SS
replist$F_std_basis
#stock0@range['minfbar'] <- 3
#stock0@range['maxfbar'] <- 7

#Define SSB time as the start of the year
m.spwn(stock0) <- 0      
harvest.spwn(stock0) <- 0

#remove forecast years
stock <- window(stock0,end=replist$endyr)

# stock summary 
plot(stock) + facet_wrap(~qname, scales="free")

```

```{r,include=F}

#check that F is ok
all(harvest(stock) ==  harvest(stock.n(stock), catch=catch.n(stock), m=m(stock)))

SSssb <- subset(replist$derived_quants,substring(Label,1,6)%in%c('SSB_19','SSB_20')) %>% 
  mutate(Yr=as.numeric(substring(Label,5,9)))
 plot(SSssb$Yr,SSssb$Value)
lines(stock@range[[4]]:stock@range[[5]],ssb(stock))
# lines(stock0@range[[4]]:stock0@range[[5]],ssb(stock0)[,,1],col='red')
legend('topright',c('SS ssb','flr ssb'),pch=c(1,NA),lty=c(NA,1),col=c(1,1))

# check recruitment
SSrec <- subset(replist$derived_quants,substring(Label,1,7)%in%c('Recr_19','Recr_20')) %>%
  mutate(Yr=as.numeric(substring(Label,6,10)))
plot(SSrec$Yr,SSrec$Value, ylim=c(0,max(SSrec$Value)))
lines(stock@range[[4]]:stock@range[[5]],rec(stock))
legend('topleft',c('SS rec','flr rec'),pch=c(1,NA),lty=c(NA,1),col=1)

# check F
SSf <- subset(replist$derived_quants,substring(Label,1,4)%in%c('F_19','F_20')) %>%
  mutate(Yr=as.numeric(substring(Label,3,7)))
plot(SSf$Yr,SSf$Value, ylim=c(0,max(SSf$Value)))
lines(stock@range[[4]]:stock@range[[5]],fbar(stock))
legend('topleft',c('SS F','FLR F'),pch=c(1,NA),lty=c(NA,1),col=1)

```


## Stock-recruit
Now get the stock-recruit para from SS and force the SR function to be the same. 
This fixes the steepness, R0 and sigma R.

```{r}
# SR para from SS
r0_null <- replist$parameters$Value[replist$parameters$Label=="SR_LN(R0)"]

regime <- exp(replist$parameters$Value[substring(replist$parameters$Label,1,13)=="SR_regime_BLK"])

R0 <- exp(r0_null)#*regime
h <- replist$parameters$Value[replist$parameters$Label=="SR_BH_steep"]


B0_null <- c(replist$derived_quants$Value[replist$derived_quants$Label=="SSB_Virgin"])
#b0 scales linearly with R0
B0 <- B0_null #* regime

# other SR para 
#sigmaR <- replist$parameters$Value[replist$parameters$Label=="SR_sigmaR"]
# better:
sigmaR <- replist$sigma_R_info[1,'SD_of_devs']

#rho <- replist$parameters$Value[replist$parameters$Label=="SR_autocorr"]
# bettter
recruit = replist$recruit
recyrs = recruit$Yr[recruit$era =="Main"]
maindevs = recruit[recruit$Yr%in%recyrs,'dev']
rho = acf(maindevs)$acf[2]

# need to calculate this for the current regime
Bmsy <- c(replist$derived_quants$Value[replist$derived_quants$Label=="SSB_MSY"])

FixedBevholt <- function (ab,ssb) log(4*h*R0*ssb/(B0*(1-h)+ssb*(5*h-1)))
# be careful to not modify h, R0 and B0 later in the script

plot(c(ssb(stock)),c(rec(stock)),xlim=c(0,4e4),ylim=c(0,1.2e5))
s <- seq(0,4e4,length=1000)
lines(s,exp((FixedBevholt(NA,s))))
lines(s,4*h*exp(r0_null)*s/(B0_null*(1-h)+s*(5*h-1)),lty=3)
legend('topleft',c('early SR','regime offset'),lty=c(3,1))
```

We believe that recruitment since 2014 is substantially lower than expected from the 
overall SR relationship from SS, so we estimate a 'regime' parameter that essentially offsets R0.
SS estimates this parameter to be `r regime/100`%. 

## Identifying appropriate data
The last 2 years the estimates of recruitment are not estimated in the model (the fish
do not show up in the catches until at least age 2). Before 2014 we assume a different regime so we 
also need to exclude the years before 2014.

In this case the year range only matters for the decision on Blim. We are not actually fitting a SR 
relationship here; all parameters are fixed here by the parameters provided by SS.

```{r}
#head(SSrec, 30)[,1:3]
#r <- tail(SSrec, 30)[,1:3]
#r %>% mutate(cv=StdDev/Value) %>% ggplot(aes(Label,cv)) + geom_point() + 
#  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

yrs <- 2014:2021
stock1 <- window(stock,start=yrs[1],end=max(yrs)) 
fit_bh <- eqsr_fit(stock1, nsamp = 1000, models = c("FixedBevholt")) #increase nsamp later
#eqsr_plot(fit_bh) # note this is without bias correction, which is correct in this context
fit_bh$sr.sto$cv <- sigmaR # force sigmaR to be consistent with ss
eqsr_plot(fit_bh) 

```

The plot above shows the SR fit with all parameters fixed to those provided by SS. So the fit
is not to the points shown here,

## Step 2. Stock type and step 3. Blim


```{r,message=F,warning=F}
fit_temp <- eqsr_fit(window(stock,start=2014,end=2021), nsamp = 1, models = c("Segreg"))
#eqsr_plot(fit_temp)
BlimSegreg <- fit_temp$sr.det$b
Blim15B0 <- .15*B0
Blim25B0 <- .25*B0
Blim15Bvirg <- .15*B0_null 
Blim2020 <- round(c(ssb(stock)[,'2020']))

nsr <- ab(fmle(as.FLSR(stock1, model='bevholtSV'),
  fixed=list(s=h, v=B0,
  spr0=B0/R0)))
brp_single <- FLBRP::brp(FLBRP::FLBRP(stock1, sr=nsr))
Bmsy_regime <- c(brp_single@refpts['msy','ssb'])

mvn = FLRef::ssmvln(replist, Fref="Btgt",verbose=F)#,years=yrs)
ts = FLRef::ss2ices(mvn)
ts1 <- subset(ts,year %in% yrs)

#average of 3 lowest above median recr
BlimEmp <- mean(sort(ts1$SSB[ts1$Rec>median(ts1$Rec)])[1:3])

b <- rbind(data.frame(name='Bmsy_early',value=Bmsy),
           data.frame(name='Bmsy_late',value=Bmsy_regime),
           data.frame(name='15%B0_late',value=Blim15B0),
#           data.frame(name='25%B0_late',value=Blim25B0),
           data.frame(name='15%B0_early',value=Blim15Bvirg),
#           data.frame(name='Segreg',value=BlimSegreg),
#           data.frame(name='Emp',value=BlimEmp)
           data.frame(name='B2020',value=Blim2020))

#ggplot(ts,aes(year,SSB)) + geom_point() +
#  geom_hline(aes(yintercept=value,col=name),b,show.legend=F,lty=2,lwd=.75) +
#  geom_errorbar(aes(ymin=SSB_low,ymax=SSB_high),col='grey',lwd=0.75) + 
#  ggrepel::geom_label_repel(aes(1990,value,label=paste(name,'\n',round(value)),col=name),b,show.legend=F,size=3) +
#  xlim(1990,NA) + ylim(0,35000)


ggplot(ts,aes(SSB,Rec)) + 
  geom_vline(aes(xintercept=value,col=name),b,show.legend=F,lty=2,lwd=.75) +
  geom_errorbar(aes(xmin=SSB_low,xmax=SSB_high),col='grey',lwd=0.75) + 
  geom_errorbar(aes(ymin=Rec_low,ymax=Rec_high),col='grey',lwd=0.75) + 
  geom_point() + geom_label(aes(label=substring(year,3,4)),size=3,col=ifelse(ts$year>=2014&ts$year<2022,'black','grey')) +
  ggrepel::geom_label_repel(aes(value,0,label=paste(name,'\n',round(value)),col=name),b,show.legend=F,size=3) +
  xlim(0,32000)+ ylim(0,130000) 
```

The plot above shows the SR pairs (the numbers refer to years; only black numbers are considered 
here). The vertical lines indicate various biomass points that may inform the choice of Blim.

The SR type cannot be classified as any of the types in the ICES guidelines. WGBSS3 considered
the following:

* Recruitment levels decreased when the stock was relatively high. Subsequently, low recruitment
then lead to a low stock size. 
* The last year with estimated recruitment deviations (2021) had the lowest
'observed' recruitment but this value is quite uncertain as the model does not have
much information on 2-year old fish.
* The most appropriate candidate for Blim may be a fraction of B0. Under the low-recruitment
regime, B0 is `r round(B0)`. For gadoids, the mean Blim/B0 is around 15% (WKNEWREF). 15% of B0
would lead to Blim below Bloss and therefore not appropriate.
* A pragmatic choice is ssb(2020). This is the highest recruitment in the recent period
and is almost identical to 15% of B0 without the reduced recruitment; so it is robust to the
assumption of low recruitment. It is also close to the estimated Bmsy under the low recruitment
regime so any higher value would be inconsistent with managing at MSY.

**Decision: Blim = B2020**

```{r}
Blim <- Blim2020
```

## step 4. Other PA reference points from Blim
```{r}
sigmaSSB_SS <- round(with(subset(replist$derived_quants,Label=='SSB_2023'),StdDev/Value),3)

SigmaSSB <- 0.2 # default

Bpa <- round(Blim * exp(1.645 * SigmaSSB))

```
SigmaSS is estimated by SS as `r round(sigmaSSB_SS,2)`, it is likely that this is an underestimate 
so we use 0.2 (default value). Bpa is therefore estimated as 1.3896 x Blim = `r round(Bpa)`

## Flim

No longer needed

## Step 2. Fmsy and Btrigger
FMSY should initially be calculated based on an evaluation with the inclusion of stochasticity 
in a population (i.e. recruitment, M, maturity, growth) and fishery (e.g. selectivity) as well 
as assessment/advice error. This is a constant F, which should provide maximum yield without 
biomass constraints (without MSY Btrigger). Error is included as this is the condition analogous 
to management strategy evaluations (MSEs) that will prevail in practice. Note that in order to 
ensure consistency between the precautionary and the MSY frameworks, FMSY is not allowed to be 
above Fpa; therefore, if the FMSY value calculated initially is above Fpa, FMSY is reduced to Fpa.

```{r}
## Forecast error (see Guidance document for details on calculation of these values)
# F
cvF  <- 0.212                                 # Default = 0.212
phiF <-	0.423                                 # Default = 0.423
# SSB
cvSSB <- 0                                    # Default = 0
phiSSB <- 0                                   # Default = 0


sim_noTrig <- eqsim_run(fit_bh,
                            Fcv=cvF, Fphi=phiF, SSBcv=cvSSB,
                            rhologRec=rho,
                            Btrigger = 0, Blim=Blim, Bpa=Bpa,
                            Nrun=200, Fscan = seq(0,1,len=101),verbose=F)

Fmsy_tmp <- round(sim_noTrig$Refs2["lanF","medianMSY"],3)
eqsim_plot_range(sim_noTrig, type="median")
```

Fmsy is initially estimated as `r Fmsy_tmp`. Note that SS estimates Fmsy at 
`r round(subset(replist$derived_quants,Label== "annF_MSY")$Value,3)`.

MSY Btrigger should be selected to safeguard against an undesirable or unexpected low 
SSB when fishing at FMSY. For most stocks that lack data on fishing at FMSY, MSY Btrigger 
is set at Bpa. However, as a stock starts to be fished consistently with FMSY, it is possible 
to move towards implementation of a value for MSY Btrigger that reflects the 5th percentile 
definition of MSY Btrigger. Here we explore the value of Bmsy5pc but not with the intention
as a candidate for Btriggger.

```{r}
data.05<-sim_noTrig$rbp
x.05 <- data.05[data.05$variable == "Spawning stock biomass", ]$Ftarget
b.05 <- data.05[data.05$variable == "Spawning stock biomass", ]$p05
plot(b.05~x.05, ylab="SSB", xlab="F")
abline(v=Fmsy_tmp)
i <- which(x.05<(Fmsy_tmp+0.1))
b.lm <- loess(b.05[i] ~ x.05[i])
lines(x.05[i],c(predict(b.lm)),type='l')

Bmsy5pc <- round(predict(b.lm,Fmsy_tmp))
abline(h=Bmsy5pc)
```

We retain Btrigger = Bpa =`r formatC(Bpa,format='d',big.mark=",")`. 
Bmsy5pc = `r  formatC(Bmsy5pc,format='d',big.mark=",")`. This is below Blim and likely
not accurately estimated.

```{r}
Btrigger <- max(Bpa, Bmsy5pc)
```

The ICES MSY AR should be evaluated to check that the FMSY and MSY Btrigger combination fulfills 
the precautionary criterion of having less than 5% annual probability of SSB < Blim in the long 
term. The evaluation must include realistic assessment/advice error and stochasticity in population 
biology and fishery selectivity.

```{r}
sim_Trig <- eqsim_run(fit_bh,
                          Fcv=cvF, Fphi=phiF, SSBcv=cvSSB,
                          rhologRec=rho,
                          Btrigger = Btrigger, Blim=Blim,Bpa=Bpa,
                          Nrun=200, Fscan = seq(0,1,len=101),verbose=F)
eqsim_plot(sim_Trig,catch=F)
eqsim_plot_range(sim_Trig, type="median")
Fp05 <- round(sim_Trig$Refs2["catF","F05"],3)
```

Fpa is estimated with the Btrigger advice rule, resulting in a much flatter yield at high F. 
Nevertheless, Fpa is estimated at `r Fp05` and therefore still limiting Fmsy (and FmsyUpper).


```{r}
# note that we compare Fp05 to the 'original' FMSY estimate, not the one with the AR and Btrigger
Fmsy <- round(min(sim_noTrig$Refs2["lanF","medianMSY"], Fp05),3)
Fupper <- round(min(sim_noTrig$Refs2["lanF","Medupper"], Fp05),3)
Flower <- round(sim_noTrig$Refs2["lanF","Medlower"],3)

if(Fmsy==Fp05) {
lan <-   data.table::data.table(sim_noTrig$rbp)[abs(Ftarget - Fmsy) == min(abs(Ftarget - Fmsy)) & variable == 'Landings', p50]
Flower <- data.table::data.table(sim_noTrig$rbp)[variable == 'Landings' & p50 >= lan * 0.95, Ftarget][1]
}
```

The estimated reference points are shown below. 


Reference Point|Value                              |Rationale
---------------|-----------------------------------|-----------------------------------------
Blim           |`r formatC(Blim,format='d',big.mark=",")`     |SSB in 2020
Bpa            |`r formatC(Bpa,format='d',big.mark=",")`      |Blim with assessment error
MSY Btrigger   |`r formatC(Btrigger,format='d',big.mark=",")` |Bpa
Fpa            |`r Fp05`                                      |F with 95% probability of SSB>Blim (BH with Btrigger)
Fmsy           |`r Fmsy`                                      |Fpa
FmsyLower      |`r Flower`                                    |Stochastic simulations
FmsyUpper      |`r Fupper`                                    |Fpa
Bmsy5pc        |`r formatC(Bmsy5pc,format='d',big.mark=",")`  |5% probability of SSB < Blim

